---
title: 'Bellabeat Exploratory Analysis'
author: "Alexandra Lazareva"
date: "2026-05-05"
output:
  pdf_document: default
  html_document: default
---

## Introduction
The goal of this case study is to analyze smart device fitness data to unlock new growth opportunities for the company. By analyzing how consumers use non-Bellabeat smart devices (specifically Fitbit data), we aim to:

Identify key trends in smart device usage.

Analyze how these trends apply to Bellabeat customers.

Provide high-level recommendations for Bellabeat’s marketing strategy based on these insights.

## Analysis and Visualization

In this section, we transition from data cleaning to exploring relationships between key variables. Our primary focus is the correlation between daily movement and energy expenditure.

### Relationship between Steps and Calories
We want to test the hypothesis that higher step counts significantly increase calorie burn, which is a core value proposition for Bellabeat's tracking features.

```{r analysis_viz, message=FALSE, warning=FALSE}
# Load necessary libraries
library(tidyverse)
library(ggplot2)

# Load the cleaned data (ensure the file is in your project folder)
daily_activity <- read.csv("daily_activity.csv")

# Create a scatter plot with a trend line
ggplot(data = daily_activity, aes(x = TotalSteps, y = Calories)) +
  geom_point(aes(color = TotalSteps), alpha = 0.5) + # Add transparency to handle overplotting
  geom_smooth(method = "lm", color = "coral") +     # Linear regression line
  scale_color_gradient(low = "blue", high = "red") + # Visualizing intensity
  labs(
    title = "Correlation: Daily Steps vs. Calories Burned",
    subtitle = "A strong positive correlation supports the drive for increased daily activity",
    x = "Total Steps",
    y = "Calories Burned",
    caption = "Data Source: Fitbit Fitness Tracker Data"
  ) +
  theme_minimal()
```

### User Activity Categorization
To provide better marketing recommendations, we categorize users based on their average activity levels. This helps Bellabeat target specific segments with personalized notifications.

```{r user_segments}
# Categorizing users by average steps
user_segments <- daily_activity %>%
  group_by(Id) %>%
  summarise(avg_steps = mean(TotalSteps)) %>%
  mutate(user_type = case_when(
    avg_steps < 5000 ~ "Sedentary",
    avg_steps >= 5000 & avg_steps < 10000 ~ "Active",
    avg_steps >= 10000 ~ "Very Active"
  ))

# Displaying the first few rows of our new table
head(user_segments)

# Calculating the Pearson correlation between steps and calories
cor_result <- cor(daily_activity$TotalSteps, daily_activity$Calories)
print(cor_result)
```


### Key Findings
1. **Positive Correlation:** As expected, there is a clear positive linear relationship between steps and calories.
2. **Engagement Gap:** A significant portion of the data points cluster below 7,500 steps, suggesting that many users struggle to reach the "Active" threshold.